TODO LIST

### 2019-02-28 LMCS/sendrev/example-3dnf-minusone.tex
sc

Let $\R = \{P_0(\cdot,\cdot,\cdot), P_1(\cdot,\cdot,\cdot), P_2(\cdot,\cdot,\cdot), P_3(\cdot,\cdot,\cdot)\}$. This vocabulary is used as follows to encode a 3-DNF formula. Relations $P_0,P_1,P_2,P_3$ evaluated on a tuple $(x,y,z)$ hold if the disjuncts $(x\wedge y \wedge z),(x\wedge y \wedge \neg z),(x\wedge \neg y \wedge \neg z),(\neg x\wedge\neg y \wedge\neg z)$ are present in the formula, respectively.
To define $\ctdnf$, we consider an $\so$-formula $\varphi(T)$ over $\R$, where $T$ is an unary predicate, such that for every 3-DNF formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sa{T}\varphi(T)$. Particularly, $T(x)$ holds if, and only if, $x$ is assigned value true. In detail,
\begin{align*}
\varphi(T) = \ex{x}\ex{y}\ex{z}\big(\,
&(P_0(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \ \; T(z))\,\vee\\
&(P_1(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \!\neg T(z))\,\vee\\
&(P_2(x,y,z)\wedge \ \; T(x)\wedge\!\neg T(y)\wedge \!\neg T(z))\,\vee\\
&(P_3(x,y,z)\wedge \!\neg T(x)\wedge \!\neg T(y)\wedge \!\neg T(z))\,\big).
\end{align*}
This formula is in $\loge{1}$, so $\ctdnf$ is in $\eqso(\loge{1})$, and indeed, also in $\eqso(\logex{1})$.

We will apply a simplified version of our construction to $\varphi$ to illustrate how it restricts a single assignment to $T$. First, note that we assumed the formula had the form $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\bar{x}$ was not empty, so we start by doing the corresponding replacement. Consider the formula is $\varphi(T,u)$ and in each disjunct the expression $\wedge \first(u)$ is added. The next step of our construction was rewriting each disjunct so a sequence of conditions (that we called {\em a, b} and {\em c}) held. After the rewritings have been done, one of the disjuncts will be
\[
\varphi'_{\ell}(T,u,x,y,z) = \underbrace{P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)}_{\varphi^{\fo}_{\ell}(u,x,y,z)}\wedge \underbrace{T(x)\wedge T(y)\wedge T(z)}_{\varphi^{+}_{\ell}(T,x,y,z)},
\]
Where $\theta(u,x,y,z)$ is a weak ordering for $(u,x,y,z)$. Furthermore,
\begin{multline*}
\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z) = P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)\, \wedge \\ \fa{u'}\fa{x'}\fa{y'}\fa{z'}\big(P_0(x',y',z')\wedge \theta(u',x',y',z')\wedge \first(u')\to(u\leq u' \wedge \langle x,y,z\rangle\leq \langle x',y',z'\rangle)\big),
\end{multline*}
and lastly,
\begin{align*}
\psi_{\ell}(T,u) = \; &\ex{u'}\ex{x}\ex{y}\ex{z}(P_0(x,y,z)\wedge \theta(u',x,y,z)\wedge \first(u'))\,\to \\
&\bigg(\ex{x}\ex{y}\ex{z}\Big[\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)\,\wedge \\ 
&\hspace{5em}\big(\,T(x)\wedge T(y)\wedge T(z)\to\ex{w}(T(w)\wedge w \neq x \wedge w\neq y \wedge w\neq z)\big)\Big]\, \vee\\
&\ex{u'}\ex{x}\ex{y}\ex{z}(\varphi'_{\ell}(T,u',x,y,z)\wedge u' < u)\bigg).
\end{align*}
Note that we omitted the $\varphi^{\fo}_{\ell}(u,x,y,z)$ part around the beginning of the third line of the formula because it is redundant. The crux of our construction are the second and third lines in the formula. Given a first-order assignment $v$ that satisfies $\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)$, consider the tuple $\langle v(x),v(y),v(z)\rangle$. We end our example by bringing attention to the fact that this tuple is a logarithmic-sized witness which effectively identifies, and lets us restrict our desired assignment to $T$.
\martin{este ejemplo sirve para ayudar a entender la idea del testigo de tama\~no logar\'itmico}

>>>>

Let $\R = \{P_0(\cdot,\cdot,\cdot), P_1(\cdot,\cdot,\cdot), P_2(\cdot,\cdot,\cdot), P_3(\cdot,\cdot,\cdot)\}$. This vocabulary is used as follows to encode a 3-DNF formula $\theta$. Relations $P_0,P_1,P_2,P_3$ evaluated on a tuple $(x,y,z)$ hold if the formulae $(x\wedge y \wedge z),(x\wedge y \wedge \neg z),(x\wedge \neg y \wedge \neg z),(\neg x\wedge\neg y \wedge\neg z)$ occur in $\theta$, respectively.
To define $\ctdnf$, we consider an $\so$-formula $\varphi(T)$ over $\R$, where $T$ is an unary predicate, such that for every 3-DNF formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sa{T}\varphi(T)$. More specifically, $T(x)$ holds if and only if  $x$ is assigned value true, so that $\varphi(T)$ is defined as follows:
\begin{align*}
\varphi(T) = \ex{x}\ex{y}\ex{z}\big(\,
&(P_0(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \ \; T(z))\,\vee\\
&(P_1(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \!\neg T(z))\,\vee\\
&(P_2(x,y,z)\wedge \ \; T(x)\wedge\!\neg T(y)\wedge \!\neg T(z))\,\vee\\
&(P_3(x,y,z)\wedge \!\neg T(x)\wedge \!\neg T(y)\wedge \!\neg T(z))\,\big).
\end{align*}
This formula is in $\loge{1}$, so $\ctdnf$ is in $\eqso(\loge{1})$, and indeed, also in $\eqso(\logex{1})$.

We apply a simplified version of our construction to $\varphi$ to illustrate how it restricts a single assignment to $T$. First, notice that we assumed the formula had the form $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\bar{x}$ was not empty, so we start by doing the corresponding replacement. Consider the formula is $\varphi(T,u)$ and in each disjunct the expression $\wedge \first(u)$ is added. The next step of our construction was rewriting each disjunct so a sequence of conditions (that we called {\em a, b} and {\em c}) held. After the rewritings have been done, one of the disjuncts will be
\[
\varphi'_{\ell}(T,u,x,y,z) = \underbrace{P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)}_{\varphi^{\fo}_{\ell}(u,x,y,z)}\wedge \underbrace{T(x)\wedge T(y)\wedge T(z)}_{\varphi^{+}_{\ell}(T,x,y,z)},
\]
Where $\theta(u,x,y,z)$ is a weak ordering for $(u,x,y,z)$. Furthermore,
\begin{multline*}
\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z) = P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)\, \wedge \\ \fa{u'}\fa{x'}\fa{y'}\fa{z'}\big(P_0(x',y',z')\wedge \theta(u',x',y',z')\wedge \first(u')\to(u\leq u' \wedge \langle x,y,z\rangle\leq \langle x',y',z'\rangle)\big),
\end{multline*}
and lastly,
\begin{align*}
\psi_{\ell}(T,u) = \; &\ex{u'}\ex{x}\ex{y}\ex{z}(P_0(x,y,z)\wedge \theta(u',x,y,z)\wedge \first(u'))\,\to \\
&\bigg(\ex{x}\ex{y}\ex{z}\Big[\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)\,\wedge \\ 
&\hspace{5em}\big(\,T(x)\wedge T(y)\wedge T(z)\to\ex{w}(T(w)\wedge w \neq x \wedge w\neq y \wedge w\neq z)\big)\Big]\, \vee\\
&\ex{u'}\ex{x}\ex{y}\ex{z}(\varphi'_{\ell}(T,u',x,y,z)\wedge u' < u)\bigg).
\end{align*}
Note that we omitted the $\varphi^{\fo}_{\ell}(u,x,y,z)$ part around the beginning of the third line of the formula because it is redundant. The crux of our construction are the second and third lines in the formula. Given a first-order assignment $v$ that satisfies $\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)$, consider the tuple $\langle v(x),v(y),v(z)\rangle$. We end our example by bringing attention to the fact that this tuple is a logarithmic-sized witness which effectively identifies, and lets us restrict our desired assignment to $T$.
\martin{este ejemplo sirve para ayudar a entender la idea del testigo de tama\~no logar\'itmico}

### 2019-02-28 LMCS/sendrev/proof-qe0-fp-qe1-totp-fptras.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
sc

\begin{lem} \label{conj-mult}
	If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
\end{lem}
\proof
	Given two formulae $\alpha, \beta$ in $\eqso(\LL)$ we will construct a formula in the logic which is equivalent to $(\alpha\mult \beta)$. 
	From what was proven in Proposition \ref{theo-pnf-snf}, we may assume that $\alpha$ and $\beta$ are in $\LL$-SNF. 
	Let $\alpha = \sum_{i = 1}^n\sa{\bar{X}_i}\sa{\bar{x}_i}\varphi_i(\bar{X}_i,\bar{x}_i)$, and $\beta = \sum_{i = 1}^m\sa{\bar{Y}_i}\sa{\bar{y}_i}\psi_i(\bar{Y}_i,\bar{y}_i)$. Expanding the product in $(\alpha\mult \beta)$ and reorganizing results in the equivalent formula
	$$
	\sum_{i = 1}^n\sum_{j = 1}^m\sa{\bar{X}_i}\sa{\bar{Y}_j}\sa{\bar{x}_i}\sa{\bar{y}_j}(\varphi_i(\bar{X}_i,\bar{x}_i)\wedge\psi_i(\bar{Y}_j,\bar{y}_j)),
	$$
	which is in $\LL$-SNF, and therefore, in $\eqso(\LL)$.
\qed
Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} can be applied to $\eqso(\loge{1})$. 
This concludes the proof.

>>>>

\begin{lem} \label{conj-mult}
	If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
\end{lem}
\proof
	Given two formulae $\alpha, \beta$ in $\eqso(\LL)$ we will construct a formula in the logic which is equivalent to $(\alpha\mult \beta)$. 
	From what was proven in Proposition \ref{theo-pnf-snf}, we may assume that $\alpha$ and $\beta$ are in $\LL$-SNF. 
	Let $\alpha = \sum_{i = 1}^n\sa{\bar{X}_i}\sa{\bar{x}_i}\varphi_i(\bar{X}_i,\bar{x}_i)$, and $\beta = \sum_{i = 1}^m\sa{\bar{Y}_i}\sa{\bar{y}_i}\psi_i(\bar{Y}_i,\bar{y}_i)$. Expanding the product in $(\alpha\mult \beta)$ and reorganizing results in the equivalent formula
	$$
	\sum_{i = 1}^n\sum_{j = 1}^m\sa{\bar{X}_i}\sa{\bar{Y}_j}\sa{\bar{x}_i}\sa{\bar{y}_j}(\varphi_i(\bar{X}_i,\bar{x}_i)\wedge\psi_i(\bar{Y}_j,\bar{y}_j)),
	$$
	which is in $\LL$-SNF, and therefore, in $\eqso(\LL)$.
\qed
Since $\loge{1}$ is closed under conjunction, we have that Lemma~\ref{conj-mult} can be applied to $\eqso(\loge{1})$, and we can deduce that $\eqso(\loge{1})$ is closed under multiplication. 
This concludes the proof of the proposition.

### 2019-02-28 LMCS/sendrev/proof-sigmafo-minusone.tex
cambios menores en -1 y agregue un ejemplo del testigo tamano llog
sc

For the sake of readability, we divide the proof in three parts. The last part, i.e. subtraction by one, is probably the most technical proof of the paper. Furthermore, since this proof is several pages long, a line was drawn on the left margin to visually differentiate the proof from the rest of the paper.
\martin{Ac\'a menciono la l\'inea del margen izquierdo}

@@

\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. 
Without loss of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$. Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ with the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.

@@

Now, we define a series of rewrites of $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$ (the precise definition of a weak ordering is detailed below) and (c) if both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.
When these conditions are met, $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ will be satisfiable if and only if $\varphi_i^{\fo}(\bar{x},\bar{y})$ is satisfiable. This is proven in Claim~\ref{claim:minusone}.
We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
\martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}

@@

The previous claim its subsequent proof motivate the following definition. For a first-order assignment $v$ over $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
One can easily check that for every assignment $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$.
Namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all satisfying assignments of $\bar{X}$ when $v$ is fixed.
This offers an insight into the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
For this, we choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
\[
\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}\big(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')\big).
\] 
If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
Furthermore, let $V^*$ be the second-order assignment and $v^*$ the first-order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
By the previous discussion, $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.

Now, we have all the ingredients in order to define $\kappa(\alpha)$. 
Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that, if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if either $V \neq V^*$ or $v \neq v^*$. 
This property can be defined as follows:
\begin{align}
\psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
&\Big( 
\ex{\bar{y}}\!\big[ \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \, \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\! \bigvee_{X \in \bar{X}}\! \ex{\bar{z}}\!( \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!\!\! \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!\! \bar{w}\neq\bar{z} \, ) \, \big)\big] \;  \vee \!  \label{eq:part2} \\ 
&\; \ex{\bar{x}'} \ex{\bar{y}}\big( \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}\big) \; \;\; \Big)  \label{eq:part3}
\end{align}
\martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exist assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
The conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), takes care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desired property.

>>>>

For the sake of readability, we divide the proof into three parts. The last part, i.e. subtraction by one, is the most technical proof in the paper. Since this proof is several pages long, a line was drawn on the left margin to visually differentiate it from the rest of the paper.
\martin{Ac\'a menciono la l\'inea del margen izquierdo}
\marcelo{OK}

@@

\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $\A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. 
Without loss of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$. Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ with the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.

@@

Now, we define a series of rewrites of $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$ (the precise definition of a weak ordering is detailed below) and (c) if both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.
When these conditions are met, $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ will be satisfiable if and only if $\varphi_i^{\fo}(\bar{x},\bar{y})$ is satisfiable. This is proven in Claim~\ref{claim:minusone}.
We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
\martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}
\marcelo{OK}

@@

The previous claim its subsequent proof motivate the following definition. For a first-order assignment $v$ over $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
One can easily check that for every assignment $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$.
Namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all satisfying assignments of $\bar{X}$ when $v$ is fixed.
This offers an insight into the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
For this, we choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
\[
\text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}\big(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')\big).
\] 
If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
Furthermore, let $V^*$ be the second-order assignment and $v^*$ the first-order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
By the previous discussion, $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.

Now, we have all the ingredients in order to define $\kappa(\alpha)$. 
Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that, if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if either $V \neq V^*$ or $v \neq v^*$. 
This property can be defined as follows:
\begin{align}
\psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}'} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x}',\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
&\Big( 
\ex{\bar{y}}\!\big[ \text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \, \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\! \bigvee_{X \in \bar{X}}\! \ex{\bar{z}}\!( \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!\!\! \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!\! \bar{w}\neq\bar{z} \, ) \, \big)\big] \;  \vee \!  \label{eq:part2} \\ 
&\; \ex{\bar{x}'} \ex{\bar{y}}\big( \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}\big) \; \;\; \Big)  \label{eq:part3}
\end{align}
\martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
\marcelo{OK}
To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exist assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
The conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), takes care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desired property.
